---
title: "Machine_Learning - Exercise_Activity_Prediction"
author: "Willianto"
date: "18/07/2020"
output: html_document
---

```{r LoadPackages, echo=FALSE}
library(knitr) ##Load Knitr package
library(ggplot2) ##Plotting and data
library(caret) ##Load package for ML
library(car) ##Load package for VIF output
```

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "hold", tidy = TRUE)
```

## R Markdown

### 1. Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. A report describing how to build the model, how to use cross validation, what are the expected out of sample error is, and why the choices are being made.

Steps Taken

1.Tidy data. Remove columns with little/no data.

2.Create Training and test data from traing data for cross validation checking

3.Trial 3 methods Random Forrest, Gradient boosted model and Linear discriminant analysis

Fine tune model through combinations of above methods, reduction of input variables or similar. The fine tuning will take into account accuracy first and speed of analysis second.


## Including Plots

